---
title: ""
output: html_document
---
## Six Month Electricty Use Forecast for Hauteville House: November 2010 - April 2011
*27 November 2010*  

Prepared For  | Prepared By
------------- | -------------:
Jean Valjean, CFO  | Jean Prouvaire, Data Scientist
Marius Pontmercy, Utilities Director  | Courfeyrac Consulting
Hauteville House, Guernsey  | Paris

<hr>

### Purpose

Describe past monthly power usage and forecast potential usage for the coming six months.

### Results

**Total monthly power use follows a predictable annual cycle, with winter peaks and summer troughs. In coming months, usage will peak between 900-1200 kWh (95% prediction range) in December and January, then steadily fall over the winter and early spring.**   

<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r everything}

# Create project directories
# Commented out here to avoid error
# dir.create("~/BIWR/Chapter1/Code", recursive=T)
# dir.create("~/BIWR/Chapter1/Data")
# dir.create("~/BIWR/Chapter1/Results")

# Set the working directory
setwd("~/BIWR/Chapter1")

# Load packages
require(zoo)
require(dplyr)
require(reshape2)
require(forecast)
require(ggplot2)
require(htmlwidgets)
require(dygraphs)

# Download the zip file into the data folder
#download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/00235/household_power_consumption.zip", destfile="Data/household_power_consumption.zip")

# Unzip the data from the zip file into the Data folder
#unzip("Data/household_power_consumption.zip", exdir="Data")

# Read the data into R 
# NAs are represented by blanks and ? in this data, so need to change to NA
power = read.table("Data/household_power_consumption.txt", sep=";", header=T, 
  na.strings=c("?",""), stringsAsFactors=FALSE)

# Convert date to an ISO date
power$Date = as.Date(power$Date, format="%d/%m/%Y")

# Create a DateTime object
power$DateTime = as.POSIXct(paste(power$Date, power$Time))

# Obtain the Month and Year for each data point
power$Month = format(power$Date,"%Y-%m")

# Add the first to each Y-m combo and convert back to ISO date
power$Month = as.Date(paste0(power$Month, "-01"))

# Use ifelse to count each minute that is NA
power$Missing = ifelse(is.na(power$Global_active_power), 1, 0)

# Use dplyr's group_by function to group the data by Date
power_group_day = group_by(power, Date)

# Use dplyr to summarize by our NA indicator 
# (where 1 = 1 minute with NA)
power_day_missing = summarize(power_group_day, Count_Missing = sum(Missing))

# Use zoo to perform interpolation for missing time series values
power$Global_active_power_locf = na.locf(power$Global_active_power)

# Use dplyr to group by month
power_group = group_by(power, Month)

# Use dplyr to get monthly max demand and total use results
power_monthly = summarize(power_group, 
  Max_Demand_kW = max(Global_active_power_locf), 
  Total_Use_kWh = sum(Global_active_power_locf)/60)

# Remove partial months from data frame
power_monthly = power_monthly[2:47,]

# Convert Month to Date
power_monthly$Month = as.Date(paste0(power_monthly$Month, "-01"))

# Create a time series object of monthly total use
total_use_ts = ts(power_monthly$Total_Use_kWh, start=c(2007,1), frequency=12)

# Automatically obtain the forecast for the next 6 months
total_use_fc = forecast(total_use_ts, h=6)

# Create a data frame with the original data
# and placeholders for the forecast details
use_df = data.frame(Total_Use = power_monthly$Total_Use_kWh, 
  Forecast = NA, Upper_80 = NA, 
  Lower_80 = NA, Upper_95 = NA, Lower_95 = NA)

# Create a data frame for the forecast details
# with a placeholder column for the original data
use_fc = data.frame(Total_Use = NA, Forecast = total_use_fc$mean, Upper_80 = 
  total_use_fc$upper[,1], Lower_80 = total_use_fc$lower[,1], 
  Upper_95 = total_use_fc$upper[,2], Lower_95 = total_use_fc$lower[,2])

# "Union" the two data frames into one
use_ts_fc = rbind(use_df, use_fc)

# Create a time series of the data and forecast results
total_use_forecast = ts(use_ts_fc, start=c(2007, 1), freq=12)

# Create the widget
energy_use_prediction_widget = dygraph(total_use_forecast, 
    main = "Predicted Monthly Electricty Use (kWh)",
    ylab = "Total kWh", width=900, height=500) %>% 
  dySeries(c("Total_Use"), label = "Actual kWh Usage") %>%
  dyEvent(x = "2008-08-01", "Closed for vacation", labelLoc = "top") %>%
  dyRangeSelector(dateWindow = c("2008-09-15", "2011-06-01")) %>%
  dyLegend(width = 800)

# Display the widget in the Viewer window
# Hit the Zoom button for a pop-out 
energy_use_prediction_widget
```

<br>

***Using the upper 80% prediction interval is a reasonable way to plan conservatively for power costs.***   

```{r forecasttable}
require(htmlTable)
total_use_fc_df = data.frame(total_use_fc)
colnames(total_use_fc_df) = c("Point Forecast", "Low 80%", "High 80%",
  "Low 95%", "High 95%")

fc_table = htmlTable(txtRound(total_use_fc_df, 0),
  col.columns = c(rep("none", 2), rep("#E6E6F0", 1), rep("none", 2)),
  css.cell = "padding-left: 3.5em; padding-right: 1em;",
  align.header="rrrrr",
  align="rrrrr")

fc_table
```

<br>  
<hr>

### Methods and Analytics Details

**Measurement Period**  

Data: 1-minute time step, `r format(min(power$Date), "%d %B %Y")` through `r format(max(power$Date), "%d %B %Y")`  
Analysis: 1-month time step, `r format(min(power_monthly$Month), "%B %Y")` through `r format(max(power_monthly$Month), "%B %Y")`  
Forecast: 1-month time step, November 2010 - April 2011  

**Data Source**  

https://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption  

**Data Dictionary**  

*Original*  

Dimensions: `r dim(power)[1]` observations of `r dim(power)[2]` variables  

1. **`date`**: date in format dd/mm/yyyy  
2. **`time`**: time in format hh:mm:ss  
3. **`global_active_power`**: household global minute-averaged active power (in kilowatts)  
4. **`global_reactive_power`**: household global minute-averaged reactive power (in kilowatts)  
5. **`voltage`**: minute-averaged voltage (in volts)  
6. **`global_intensity`**: household global minute-averaged current intensity (in amperes)   
7. **`sub_metering_1`**: energy sub-metering No. 1 (in watt-hours of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven, and a microwave (hot plates are not electric but gas powered).  
8. **`sub_metering_2`**: energy sub-metering No. 2 (in watt-hours of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator, and a light.  
9. **`sub_metering_3`**: energy sub-metering No. 3 (in watt-hours of active energy). It corresponds to an electric water-heater and an air-conditioner.   

**`global_active_power`** is the variable of interest.  

Missing values pattern:  

```{r missingplot}
# Download the 'calendarHeat' function from revolutionanalytics.com
source("http://blog.revolutionanalytics.com/downloads/calendarHeat.R")

# Plot the calendar graph to view the missing data pattern
calendarHeat(power_day_missing$Date, power_day_missing$Count_Missing, 
    varname="Missing Data", color="w2b")
```

*Transformations*  

Missing values were interpolated using the 'carry last value forward' approach. Comparison of distributions:  

```{r missingvalues, fig.height=2.5}
# Compare the original and interpolated distributions
# Reshape the two variables into long form for ggplot
power_long = melt(power, id.vars= "DateTime", measure.vars=
    c("Global_active_power", "Global_active_power_locf"))

# Create density plot
density_plot = ggplot(power_long, aes(value, fill=variable, 
    color=variable)) +
  geom_density(alpha=0.75) +
  facet_wrap(~variable)

# Display plot
density_plot
```

Values were aggregated to a monthly time step.   

```{r monthlyuse, fig.height=2.5}
# Create plot of total use by month
total_use_plot = ggplot(power_monthly, aes(Month, Total_Use_kWh)) +
    geom_line(col="blue", lwd=1) 

# Display plot
total_use_plot
```

*Final*  

Dimensions: `r dim(power_monthly)[1]` observations of `r dim(power_monthly)[2]` variables  

R time series object  
1. **`Month`**: Observation or forecast month (`ts` identifier, not a variable in the object)  
2. **`Total_Use`**: Total monthly active power use for period of record (in kilowatt-hours)  
3. **`Forecast`**: Point forecast of total monthly active power use (in kilowatt-hours)  
4. **`Upper_80`**: Upper 80% prediction interval for the forecast of total monthly active power use (in kilowatt-hours)  
5. **`Lower_80`**: Lower 80% prediction interval for the forecast of total monthly active power use (in kilowatt-hours)  
6. **`Upper_95`**: Upper 95% prediction interval for the forecast of total monthly active power use (in kilowatt-hours)  
7. **`Lower_95`**: Lower 95% prediction interval for the forecast of total monthly active power use (in kilowatt-hours)  

```{r basic_fc_plot, fig.height=3}
# View model summary
summary(total_use_fc)

# View the forecast plot
plot(total_use_fc)
```

***Session Info***

File compiled: 
```{r sessioninfo, echo=FALSE, warning=FALSE, message=FALSE}
Sys.time()
sessionInfo()
```

***End of File***
